An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets

### 2021-09-01 CMPSpaper_writing/CMPSpaper_writing.Rmd
small edits

2. **identify candidate positions:** For each basis segment a set of candidate registration  positions on the comparison signature is identified based on the segment's similarity to the reference signature. In a first step, the cross-correlation function of the segment to the reference is calculated, then a number of positions with high correlation values are identified as candidate positions. In case multiple segment lengths are considered, the length of each basis segment is doubled \wj{by default} and these two steps are repeated. Only when candidate positions coincide (or are similar enough), they are considered further. \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.

@@

        -   If `npeaks.set` is an integer vector of length more than 1, for example `npeaks.set = c(5, 3, 1)`, the multi-segment lengths strategy will be used: \wj{calculate the cross-correlation function between a basis segment and the reference and identify positions of the top five peaks; double or increase the legment length to a specified value, recalculate the cross-correlation function, and identify three peaks; repeat this process and identify one peak in the newly computed cross-correlation function.} \autoref{fig:seg_all} shows an example of three levels of basis segment 6 and their corresponding cross-correlation curves and identified peaks. \wj{Note that in @cmps the segment length is doubled at each level of a basis segment, but users are allowed to choose the segment length at each level in our implementation.}

@@

    +  **multi-segment lengths strategy:** with the multi-segment lengths strategy being used, a position is identified as a candidate position for registration and is called as a "consistent correlation peak" if it results in a top peak in the cross-correlation curve with a tolerance zone determined by `Tx` in all segment levels. Note that in @cmps, a segment at its largest scale (highest level) always identifies one peak \wj{, but we do not have this requirement in our implementation}.
        -   the function `get_seg_scale(segments, nseg, out_length)` is used to obtain the \wj{increased} version of a basis segment. `segments`, which is a list containing all basis segments generated by the function `get_segs(...)` in step 1, and `nseg` are used to determine the basis segment to be increased. `out_length` \wj{specifies the length of the output segment.}

@@

* `outlength` specifies the segment length of a basis segment at each level under the multi-segment lengths strategy. By default `outlength` is set to `NULL`, \wj{indicating that a basis segment should double its segment length at the next level and conforming to the description in} @cmps .

@@

\wj{The CMPS score can be considered as a similarity score of a land-level comparison, involving two bullet signatures generated from two bullet land engravings. Since each bullet has 6 land engravings in our case, a comparison of two bullets involves 36 land-level comparisons, resulting in 36 CMPS scores (as shown in \autoref{fig:tiles}). In order to obtain a similarity score of a bullet-level comparison summarizing these 36 CMPS scores,} researchers introduced two similarity \wj{metrics (or measures? XXX)}, $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, for bullet-level comparisons [@cmps]. $\mathrm{CMPS_{max}}$ is the highest CMPS score obtained among all land-level comparisons. And $\mathrm{\overline{CMPS}_{max}}$ is the highest possible mean CMPS score of land-level comparisons that are in the same phase.

>>>>

2. **identify candidate positions:** For each basis segment a set of candidate registration  positions on the comparison signature is identified based on the segment's similarity to the reference signature. In a first step, the cross-correlation function of the segment to the reference is calculated, then a number of positions with high correlation values are identified as candidate positions. In case multiple segment lengths are considered, the length of each basis segment is expanded (by default it is doubled) and these two steps are repeated. Only when candidate positions coincide (or are similar enough), they are considered further. \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.

@@

        -   If `npeaks.set` is an integer vector of length more than 1, for example `npeaks.set = c(5, 3, 1)`, the multi-segment lengths strategy will be used: calculate the cross-correlation function between a basis segment and the reference and identify positions of the top five peaks; double the segment length to a specified value, re-calculate the cross-correlation function, and identify three peaks; repeat this process and identify a single peak in the newly computed cross-correlation function. \autoref{fig:seg_all} shows an example of three levels of basis segment 6 and their corresponding cross-correlation curves and identified peaks. Note that in @cmps the segment length is doubled at each level of a basis segment, but in the present implementation users are allowed to choose the segment length at each level.

@@

    +  **multi-segment lengths strategy:** with the multi-segment lengths strategy being used, a position is identified as a candidate position for registration and is called as a "consistent correlation peak" if it results in a top peak in the cross-correlation curve with a tolerance zone determined by `Tx` in all segment levels. Note that in @cmps, a segment at its largest scale (highest level) always identifies one peak, but we do not have this requirement in our implementation.
        -   the function `get_seg_scale(segments, nseg, out_length)` is used to obtain the (potentially increased) version of a basis segment. `segments`, which is a list containing all basis segments generated by the function `get_segs(...)` in step 1, and `nseg` are used to determine the basis segment to be increased. `out_length` specifies the length of the output segment.

@@

* `outlength` specifies the segment length of a basis segment at each level under the multi-segment lengths strategy. By default `outlength` is set to `NULL`, indicating that a basis segment should double its segment length at the next level and conforming to the description in @cmps .

@@

The CMPS algorithm measures the similarity between two signatures resulting in a similarity score of a land-to-level comparison. Bullets fired from traditionally rifled barrels have multiple land and groove engraved areas. Here, we are working with bullets fired from Ruger P85 barrels with six lands and grooves. A comparison of two bullets therefore involves 36 land-to-land comparisons, resulting in 36 CMPS scores (as shown in \autoref{fig:tiles}). In order to obtain a single similarity score of a bullet-level comparison, we need to summarize these 36 CMPS scores. Two similarity metrics for bullet-level comparisons have been introduced in the literature [@cmps]: $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$. $\mathrm{CMPS_{max}}$ is the highest CMPS score obtained among all land-level comparisons, while $\mathrm{\overline{CMPS}_{max}}$ is the highest possible mean CMPS score of land-level comparisons that are in the same phase:
